---
title: "01_load"
format: html
editor: visual
---

# Load Data

```{r}
#| warning: false
library("tidyverse")
library("readxl")
```

## Loading data and creating directories

Our data comes from NCBI and is in an excel format. We use the URL and readxl to load the data.

```{r}
#| label: load data and directories

# Creating directories
data_dir <- "../data/"
raw_dir <- "../data/_raw/"
doc_dir <- "../doc/"
results_dir <-"../results/"

if (!dir.exists(data_dir)){
  dir.create(path = data_dir)
}
if (!dir.exists(raw_dir)){
  dir.create(path = raw_dir)
}
if (!dir.exists(doc_dir)){
  dir.create(path = doc_dir)
}
if (!dir.exists(results_dir)){
  dir.create(path = results_dir)
}


# Data retrieval
data_file <- "GSE182951_Counts_All_Samples.xlsx"

data_loc <- "https://www.ncbi.nlm.nih.gov/geo/download/?acc=GSE182951&format=file&file=GSE182951%5FCounts%5FAll%5FSamples%2Exlsx"

if (!file.exists(str_c(raw_dir, data_file))){
  download.file(
    url = str_c(data_loc),
    destfile = str_c(raw_dir, data_file))
  }

reads_data = read_excel(str_c(raw_dir, data_file))
```

## Creating the metadata

Based on the paper we know that:

1.  Each sample is paired: 1 - 2, 3 - 4, 5 - 6, etc.
2.  Each pair is split into pre-treatment (uneven numbers) and post-treatment (even numbers).
3.  Samples: 1-12 belong to group A - pCR, the rest belong to group B - no pCR.

```{r}

meta_data <- data.frame(ID = 1:40) |> 
  
  #Adding pre/post treatment to the dataset
  mutate(Treatment = case_when(str_ends(ID,"[13579]") ~ "pre_treatment",
                               str_ends(ID,"[^13579]") ~ "post_treatment")) |>
  
  # Adding corresponding Person ID
  mutate(subject_ID = case_when(Treatment == "pre_treatment" ~ (ID+1)/2,
                                Treatment == "post_treatment" ~ ID/2)) |> 
  # Changing subject ID number
  mutate(subject_ID = case_when(subject_ID < 10 ~ str_c("subject_0", subject_ID),
                                subject_ID >= 10 ~ str_c("subject_", subject_ID))) |> 
  #Adding pCR status
  mutate(pCR_status = case_when(ID <= 12 ~ "pCR",
                                ID > 12 ~ "no_pCR"))

write_csv(x = meta_data,
          file = "../data/meta_data.csv")

```
