---
title: "02_clean"
format: html
editor: visual
---

```{r}
library(tidyverse)
```

```{r}
# First we explore the data
#View(expression_data)
expression_data |> 
  sample_n(10)
```

Column:

**Name** which include the gene names.

**Identifier** which includes a link to the ENSG identifier for the gene.

Then 40 columns for each sample. They start with 20756_0XX_genes - Total counts. Where XX is the number of the sample

So we need to clean the data and change it into tidy data meaning:

Columns = variables

Rows = Observations

Each cell = values.

**Cleaning:**

-   Create tidy data

-   Clean up names

```{r}
expression_data_long <- expression_data |> 
  
  #Pivot longer to make into tidy data set.
  pivot_longer(cols = starts_with("20756"),
               names_to = "Sample",
               values_to = "Count") |>
  
  # Change to sample number
  mutate(Sample = as.numeric(str_sub(Sample, 
                                     start = 7, 
                                     end = 9))) |>  
  
  #Rename Name column 
  rename("Name" = "Gene")

expression_data_long
```

# Joining meta data and expression data

```{r}
#| label: Metadata added


# Connect the expression dataset Sample with the metadata set ID

full_expression_data <- inner_join(x = expression_data_long, 
                                   y = meta_data, 
                                   join_by(Sample == ID))

write_csv(full_expression_data, 
          "../data/full_expression_data.csv")

```
